---
format: gfm
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
---

<!-- README.md is generated from README.qmd. Please edit that file -->

# ksjdata

<!-- badges: start -->
[![R-CMD-check](https://github.com/UchidaMizuki/ksjdata/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/UchidaMizuki/ksjdata/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

English | [日本語](README.ja.md)

ksjdata provides National Land Numerical Information (国土数値情報, KSJ), published by the Ministry of Land, Infrastructure, Transport and Tourism of Japan.

- `ksj_available` is the catalog: it lists every supported file, with its dataset, release, area, and terms of use. Filter it to choose files.
- `ksj_get()` downloads and reads one file, and returns one layer as an `sf` tibble. To get several files, iterate over their names, for example with `purrr::map()`. Each file is cached by the [tarchives](https://github.com/UchidaMizuki/tarchives) package, so it is downloaded only once.
- Columns are named in Japanese after the official attribute table. Use `col_names = "raw"` to keep the names of the file (usually attribute codes such as `N03_001`), for example for older releases.

## Installation

You can install the development version of ksjdata from GitHub with:

``` r
# install.packages("pak")
pak::pak("UchidaMizuki/ksjdata")
```

## Usage

Filter `ksj_available` to find files. Each row is one file, identified by its `file_name` (the name of the archive on the KSJ download site).

```{r}
#| message: false
library(ksjdata)
library(dplyr)

files <- ksj_available |>
  filter(
    dataset_code == "N03",
    year == 2026,
    prefecture_code %in% c("13", "14")
  )
files |>
  select(dataset_name, year_name, area_name, file_name, file_size)
```

Pass a `file_name` to `ksj_get()` to get that file. The terms of use are shown the first time a dataset is requested in a session.

```{r}
ksj_get("N03-20260101_13_GML.zip")
```

To get several files, iterate over their names:

```{r}
#| eval: false
purrr::map(files$file_name, ksj_get)
```

In a [targets](https://docs.ropensci.org/targets/) pipeline, `ksj_target()` declares a target that gets one file. Use `tarchetypes::tar_eval()` to declare one per file.

```{r}
#| eval: false
tarchetypes::tar_eval(
  ksj_target(name, file_name),
  values = list(
    name = rlang::syms(stringr::str_c("n03_", files$prefecture_code)),
    file_name = files$file_name
  )
)
```

The data are returned as published: geometries and coordinate reference systems are not changed, codes keep their leading zeros, and published zeros stay distinct from blank cells (`NA`).

## Terms of use

Each dataset has its own terms, shown in the `license_*` columns of `ksj_available` and the first time the dataset is requested in a session. Some datasets cannot be used commercially (非商用). Published or processed data must credit the source, and processed data must say that they were processed.

- The terms are copied from the KSJ site as published, and may change. The pages of the site (`license_url` and `page_url`) are authoritative.
- Terms can differ by release or by area within a dataset (for example, older releases of some datasets cannot be used commercially, and some areas have their own conditions). `license_name` is the label of the dataset as a whole; read `license_note` for these conditions.
- Complying with the terms is up to you. This package does not give legal advice.
- The MIT license of this package applies to its code only, not to the data.

## Supported datasets

```{r}
#| echo: false
#| message: false
library(dplyr)

ksjdata::ksj_available |>
  summarise(
    Name = first(dataset_name),
    Releases = stringr::str_c(min(year), max(year), sep = "–"),
    Terms = first(license_name),
    .by = dataset_code
  ) |>
  arrange(dataset_code) |>
  rename(Code = dataset_code) |>
  knitr::kable()
```
