---
format: gfm
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
---

<!-- README.md is generated from README.qmd. Please edit that file -->

# ksjdata

English | [日本語](README.ja.md)

ksjdata provides National Land Numerical Information (国土数値情報, KSJ), published by the Ministry of Land, Infrastructure, Transport and Tourism of Japan.

- `ksj_available` is the catalog: it lists every supported file, with its dataset, release, area, and terms of use. Filter it to choose files.
- `ksj_get()` downloads and reads the files you choose, and returns them as one `sf` tibble. Each file is cached by the [tarchives](https://github.com/UchidaMizuki/tarchives) package, so it is downloaded only once.
- Columns are named in Japanese after the official attribute table. Use `col_names = "raw"` to keep the names of the file (usually attribute codes such as `N03_001`), for example for older releases.

## Installation

You can install the development version of ksjdata from GitHub with:

``` r
# install.packages("pak")
pak::pak("UchidaMizuki/ksjdata")
```

## Usage

```{r}
#| eval: false
library(ksjdata)
library(dplyr)

files <- ksj_available |>
  filter(dataset_code == "N03", year == 2026, prefecture_code == "13")
ksj_get(files)
```

The data are returned as published: geometries and coordinate reference systems are not changed, codes keep their leading zeros, and published zeros stay distinct from blank cells (`NA`).

## Terms of use

Each dataset has its own terms, shown in the `license_*` columns of `ksj_available` and the first time the dataset is requested in a session. Some datasets cannot be used commercially (非商用). Published or processed data must credit the source, and processed data must say that they were processed.

- The terms are copied from the KSJ site as published, and may change. The pages of the site (`license_url` and `page_url`) are authoritative.
- Terms can differ by release or by area within a dataset (for example, older releases of some datasets cannot be used commercially, and some areas have their own conditions). `license_name` is the label of the dataset as a whole; read `license_note` for these conditions.
- Complying with the terms is up to you. This package does not give legal advice.
- The MIT license of this package applies to its code only, not to the data.

## Supported datasets

```{r}
#| echo: false
#| message: false
library(dplyr)

ksjdata::ksj_available |>
  summarise(
    Name = first(dataset_name),
    Releases = stringr::str_c(min(year), max(year), sep = "–"),
    Terms = first(license_name),
    .by = dataset_code
  ) |>
  arrange(dataset_code) |>
  rename(Code = dataset_code) |>
  knitr::kable()
```
